Parsing Postfix log files

### 2008-05-01 doc/logparser-other-parsers-review.tex
Added summary table to other parsers' conclusion. Some more details about AWStats. Started testing LMA.

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline
There does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix logs, in
fact it operates by coercing Postfix logs into Apache\footnote{The Apache
web server is the most popular HTTP server in use for the past 10 years;
more information is available at \url{http://httpd.apache.org/}.} format
log files, for analysis by AWStats' HTTP log file parser.  The converting
parser only deals with a small portion of the log lines generated by
Postfix, silently skipping those it cannot deal with, and does not
distinguish between different types of rejection; there is no easy way to
extend it to handle new log lines.  Although though it does correlate log
lines by queueid, it does not deal with any of the other complications
described in this document.

When tested with the \numberOFlogFILES{} test log files AWStats' report
says it parsed 9,240,075 (88.70\%) log lines out of 10,416,129, with
1,176,050 (11.29\%) corrupt lines; however there are actually
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files.  AWStats
actually parsed 17.15\% of the input lines, ignoring the remaining 82.85\%.
The graphs it produces give an overview of mails received for the last
calendar month, showing:

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the logs; those which don't discard the data retain it
in a format inaccessible to other tools.  All of the parsers reviewed can
produce a report of greater or lesser detail and complexity, a facility not
offered by \parsername{}; reporting is deferred to a subsequent
application.

>>>>

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix logs, in
fact it operates by coercing Postfix logs into Apache\footnote{The Apache
web server is the most popular HTTP server in use for the past 10 years;
more information is available at \url{http://httpd.apache.org/}.} format
log files, for analysis by AWStats' HTTP log file parser.  The converting
parser only deals with a small portion of the log lines generated by
Postfix, silently skipping those it cannot deal with, and does not
distinguish between different types of rejection; there is no easy way to
extend it to handle new log lines.  Although though it does correlate log
lines by queueid, it does not deal with any of the other complications
described in this document.  AWStats supports saving data but the format of
the saved data is not documented (as far as the author could tell).  It
also supports reading compressed log files, but that functionality was not
tested.

When tested with the \numberOFlogFILES{} test log files AWStats' report
says it parsed 9,240,075 (88.70\%) log lines out of 10,416,129, with
1,176,050 (11.29\%) corrupt lines; however there are actually
\numberOFlogLINES{} lines in the \numberOFlogFILES{} log files, so AWStats
only parsed 17.15\% of the input lines, ignoring the remaining 82.85\%.
The graphs it produces give an overview of mails received for the last
calendar month, showing:

@@

\begin{table}[ht]
\begin{tabular}{l|l|l|l|l}
    Parser              & Parsed     & Data        & Custom        & Compressed \\
                        & test logs? & store?      & reports?      & logs       \\
    \hline                                                                      \\
    Pflogsumm           & Yes        & No          & No            & No         \\
    Sawmill             & Yes        & Yes         & Searches      & Partially  \\
    Splunk              & Yes        & Yes         & Searches      & Yes        \\
                        &            &             & \& reports    &            \\
    Isoqlog             & No         & Yes         & No            & No         \\
    AWStats             & Partially  & Yes         & Partially\dag & Yes        \\
    Throughput monitor  & N/A        & No          & No            & No         \\
    Anteater            & No         & No          & No            & No         \\
    YAALA               & No         & Yes\ddag    & Yes           & No         \\
    Logparser/Lire      & Yes        & Yes         & Yes           & Yes        \\
    Logrep              & No         & Yes         & No            & No         \\
    Log Mail Analyser   & Yes        & Yes         & Partially     & No         \\ 
    \parsername{}       & Yes        & Yes\dag\dag & No            & Yes        \\
\end{tabular}

\dag AWStats can be configured to omit a section from a report, but not to
add extra sections.

\ddag YAALA only stores the data required to produce the latest report;
other data will be discarded.

\dag\dag\parsername{} is the only parser with documentation for its data store.

\caption{Summary of parsers' features}
\label{Summary of parsers' features}
\end{table}

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the logs; those which don't discard the data retain it
in a format inaccessible to other tools.  All of the parsers reviewed can
produce a report of greater or lesser detail and complexity, a facility not
offered by \parsername{}; reporting is deferred to a subsequent
application.  Table \refwithpage{Summary of parsers' features} provides a
summary of the parsers' features.
